# Add Column using mutate
# mutate() takes a name-value pair as an argument. 
# The name will be the name of the new column you are adding, and the value is an expression defining the values 
# of the new column in terms of the existing columns mutate() returns a new data frame with the added column.

---
title: "Modifying Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
dogs <- read_csv('dogs.csv')
```

```{r}
# inspect data frame
head(dogs)
```

```{r}
# add average height column
dogs <- dogs %>%
  mutate(avg_height = (height_low_inches + height_high_inches)/2)
head(dogs)
```


# Add Multiple Columns 
---
title: "Modifying Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
dogs <- read_csv('dogs.csv')
```

```{r}
# inspect data frame
head(dogs)
```

```{r}
# add average height, average weight and rank change columns
dogs <- dogs %>%
  mutate(avg_height = (height_low_inches + height_high_inches)/2,
         avg_weight = (weight_low_lbs + weight_high_lbs)/2,
        rank_change_13_to_16 = rank_2016 - rank_2013)
head(dogs)
```


# Transmute Columns (add new columns while dropping the existing columns that may no longer be useful for your analysis.)
# Like mutate(), transmute() takes name-value pairs as arguments. 
# The names will be the names of the new columns you are adding, and the values are expressions defining the 
# values of the new columns. The difference, however, is that transmute() returns a data frame 
# with only the new columns.
---
title: "Modifying Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
dogs <- read_csv('dogs.csv')
```

```{r}
# inspect data frame
head(dogs)
```

```{r}
# update the function call to drop all existing columns
dogs <- dogs %>%
  transmute(breed = breed,
            avg_height = (height_low_inches + height_high_inches)/2,
         avg_weight = (weight_low_lbs + weight_high_lbs)/2,
        rank_change_13_to_16 = rank_2016 - rank_2013)
head(dogs)
```


# Rename Columns
# rename() can take any number of arguments, where each new column name is assigned to 
# replace an old column name in the format new_column_name = old_column_name. rename() returns a new data frame with the updated column names.
---
title: "Modifying Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
dogs <- read_csv('dogs.csv')
```

```{r message=FALSE}
# add columns and remove existing columns
dogs <- dogs %>%
  transmute(breed = breed,
            avg_height = (height_low_inches + height_high_inches)/2,
         avg_weight = (weight_low_lbs + weight_high_lbs)/2,
        rank_change_13_to_16 = rank_2016 - rank_2013)
```

```{r}
# check column names
original_col_names <- colnames(dogs)
original_col_names
```

```{r}
# rename data frame columns
dogs <- dogs %>%
  rename(avg_height_inches = avg_height,
         avg_weight_lbs = avg_weight,
         popularity_change_13_to_16 = rank_change_13_to_16)
```

```{r}
# check column names
new_col_names <- colnames(dogs)
new_col_names
```


# Add new columns, drop existing columns and arrange
---
title: "Modifying Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
dogs <- read_csv('dogs.csv')
```

```{r}
# inspect data frame
head(dogs)
```

```{r}
# add new columns, drop existing columns and arrange
dogs <- dogs %>%
  transmute(breed = breed,
            height_average_feet = ((height_low_inches + height_high_inches)/2)/12,
            popularity_change_15_to_16 = rank_2016 - rank_2015) %>% 
  arrange(desc(popularity_change_15_to_16))
head(dogs)
```



# Summary
# add new columns to a data frame using mutate()
# add new columns to a data frame and drop existing columns using transmute()
# change the column names of a data frame using rename()