# For data to be tidy, it must have:
# Each variable as a separate column
# Each row as a separate observation

# head() — display the first 6 rows of the table
# summary() — display the summary statistics of the table
# colnames() — display the column names of the table


# Handling multiple files
---
title: "Data Cleaning in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r}
# list files
student_files <- list.files(pattern = 'exams_.*csv')
print(student_files)
```

```{r message=FALSE}
# read files
# read each file in files into a data frame with read_csv(), storing the data frames in df_list
df_list <- lapply(student_files,read_csv)
```

```{r}
# concatenate data frames
students <- bind_rows(df_list)
students
```

```{r}
# number of rows in students
nrow_students <- nrow(students)
print(nrow_students)
```


# Reshape Data
# gather() takes a data frame and the columns to unpack
---
title: "Data Cleaning in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(dplyr)
library(tidyr)
```

```{r}
# load students data frame
load("students.Rda")
```

```{r}
# original column names
original_col_names <- colnames(students)
print(original_col_names)
```

```{r}
# gather columns
students <- students %>%
  gather('fractions','probability',key='exam', value='score')
head(students)
```

```{r}
# updated column names
gathered_col_names <- colnames(students)
print(gathered_col_names)
```

```{r}
# unique value counts of exam
exam_counts <- students %>%
  count(exam)
exam_counts
```


# Duplicated data
# To check for duplicates, we can use the base R function duplicated(), 
# which will return a logical vector telling us which rows are duplicate rows.

---
title: "Data Cleaning in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(dplyr)
library(tidyr)
```

```{r}
# load students data frame
load("students.Rda")
```

```{r}
# drop id column
students <- students %>%
  select(-id)
head(students)
```

```{r}
# find and count duplicated rows
duplicates <- students %>%
  duplicated() %>%
  table()
duplicates
```

```{r}
# remove duplicated rows
students <- students %>%
  distinct()
```

```{r}
# find and count duplicated rows in updated data frame
updated_duplicates <- students %>%
  duplicated() %>%
  table()
updated_duplicates
```


# Substrings str_sub()
---
title: "Data Cleaning in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(dplyr)
library(stringr)
```

```{r}
# load students data frame
load("students.Rda")
```

```{r}
# print columns of students
print(colnames(students))
```

```{r}
# view head of students
head(students)
```

```{r}
# add gender and age columns
students <- students %>%
  mutate(gender = str_sub(gender_age,1,1),
         age = str_sub(gender_age,2))
head(students)
```

```{r}
# drop gender_age column
students <- students %>%
  select(-gender_age)
head(students)
```


# Split on specific characters using separate()
---
title: "Data Cleaning in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(dplyr)
library(tidyr)
```

```{r}
# load students data frame
load("students.Rda")
```

```{r}
# view the head of students
head(students)
```

```{r}
# separate the full_name column in first_name and last_name using a space as the focus of seperation
students <- students %>%
  separate(full_name,c('first_name','last_name'),' ',extra ='merge')
head(students)
```


# Internal structure of an R object
# str() displays the internal structure of an R object. 
# Calling str() with a data frame as an argument will return a variety of information, including the data types.
---
title: "Data Cleaning in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(dplyr)
```

```{r}
# load students data frame
load("students.Rda")
```

```{r}
# print structure of students
print(str(students))
```

```{r}
# mean of age column
students %>% 
  summarise(mean_age = mean(age))
```


# String Parsing
# gsub() can use regular expressions
# as.numeric() can convert character strings containing numerical values to numeric
---
title: "Data Cleaning in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(dplyr)
```

```{r}
# load students data frame
load("students.Rda")
```

```{r}
# view head of students
head(students)
```

```{r}
# remove % from score column
students <- students %>%
  mutate(score=gsub('\\%','',score))
head(students)
```

```{r}
# change score column to numeric
students <- students %>%
  mutate(score=as.numeric(score))
head(students)
```