# Read CSV
---
title: "Introduction to Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r}
# load data frame
artists <- read_csv('artists.csv')
artists
```

# Write from Dataframe to CSV
# write_csv(artists,'new_csv_file.csv')


# Head for row numbers & Summary for row stats
---
title: "Introduction to Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
artists <- read_csv('artists.csv')
```

```{r}
# inspect data frame
head(artists)
summary(artists)
```

# Piping
# Pipe operator, or %>%, helps increase the readability of data frame code by piping the value on its left into the first argument 
# of the function that follows it. For example:

# easier to read than
df %>% 
  head()

# this
head(df)



# Select example (just like SQL select)

---
title: "Introduction to Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
artists <- read_csv('artists.csv')
```

```{r}
# select one column
artist_groups <- artists %>%
  select(group)
artist_groups
```

```{r}
# select multiple columns
group_info <- artists %>%
  select(group,spotify_monthly_listeners,year_founded)
group_info
```


# Select All Columns except some
---
title: "Introduction to Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
artists <- read_csv('artists.csv')
```

```{r}
# select all columns except one
no_albums <- artists %>% 
  select(-albums)
no_albums
```

```{r}
# select all columns except a set
df_cols_removed <- artists %>% 
  select(-genre,-spotify_monthly_listeners,-year_founded)
df_cols_removed
```


# Filter results like Where clause
---
title: "Introduction to Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
artists <- read_csv('artists.csv')
```

```{r}
# filter rows one condition
rock_groups <- artists %>%
  filter(genre == 'Rock')
rock_groups
```

```{r}
# filter rows multiple conditions
popular_rock_groups <- artists %>%
  filter(genre == 'Rock', spotify_monthly_listeners > 20000000)
popular_rock_groups
```


# Filter using Not and Or
---
title: "Introduction to Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
artists <- read_csv('artists.csv')
```

```{r}
# filter rows with or
korea_or_before_2000 <- artists %>% 
  filter(country == 'South Korea' | year_founded < 2000)
korea_or_before_2000
```

```{r}
# filter rows with not !
not_rock_groups <- artists %>% 
  filter(!(genre == 'Rock'))
not_rock_groups
```

# Order rows
---
title: "Introduction to Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
artists <- read_csv('artists.csv')
```

```{r}
# arrange rows in ascending order
group_asc <- artists %>%
  arrange(group)
group_asc
```

```{r}
# arrange rows in descending order
youtube_desc <- artists %>%
  arrange(desc(youtube_subscribers) )
youtube_desc
```

# Full tasks example 
---
title: "Introduction to Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
artists <- read_csv('artists.csv')
```

```{r}
# select columns
chosen_cols <- artists %>% 
  select(-country,-year_founded,-albums)
head(chosen_cols)
```

```{r}
# filter rows
popular_not_hip_hop <- chosen_cols %>% 
  filter(spotify_monthly_listeners > 20000000, genre != 'Hip Hop') 
head(popular_not_hip_hop)
```

```{r}
# arrange rows
youtube_desc <- popular_not_hip_hop %>% 
  arrange(desc(youtube_subscribers))
head(youtube_desc)
```

```{r}
# select columns, filter and arrange rows
artists <- artists %>% 
  select(-country,-year_founded,-albums) %>% 
  filter(spotify_monthly_listeners > 20000000, genre != 'Hip Hop') %>% 
  arrange(desc(youtube_subscribers))
head(artists)
```

# Summary
# load data from a CSV into a data frame
# inspect the data frame with head() and summary()
# select() the columns you want to analyze
# filter() the rows with comparison and logical operators
# arrange() rows in ascending or descending order
# You’ve also been exposed to the pipe %>%, a powerful tool for chaining function calls, as well as the general principles of data manipulation.