# class(2) # numeric
# class('hello') # character
# class('23') #character
# class (FALSE) #logical
# class(NA) #logical

---
title: "Introduction to R Syntax"
output: html_notebook
---
```{r}
print('Mark')
print(1000)
print('1000')

```

# Variables

---
title: "Introduction to R Syntax"
output: html_notebook
---
```{r}
name <- "Mark"
age <- 9090
```


# Vectors (yeah it's really arrays)

# You can check the type of elements in a vector by using typeof(vector_name)
# You can check the length of a vector by using length(vector_name)
# if we wanted to access the second element we would write: vector_name[2]

---
title: "Introduction to R Syntax"
output: html_notebook
---
```{r}
# We created this vector by separating 3 character strings with a comma and wrapping them inside c().
phone <- c(123,345,6789)
```


# Conditionals (control flow)
---
title: "Introduction to R Syntax"
output: html_notebook
---
```{r}
message <- 'I change based on a condition.'
if (TRUE) {
   message <- 'I execute this when true!'
} else {
   message <- 'I execute this when false!'
}

print(message)
```

# Comparison operators
# Less than: <
# Greater than: >
# Less than or equal to: <=
# Greater than or equal to: >=
# Is equal to: ==
# Is NOT equal to: !=


# Logical operators
# the and operator (&)
# the or operator (|)
---
title: "Introduction to R Syntax"
output: html_notebook
---
```{r}

message <- 'Should I pack an umbrella?'
weather <- 'cloudy'
high_chance_of_rain <- TRUE


if (weather == 'cloudy' & high_chance_of_rain) {
  message <- "Pack umbrella!"
} else {
  message <- "No need for umbrella!"
}

print(message)

```


# Function calling
---
title: "Introduction to R Syntax"
output: html_notebook
---
```{r}
data <- c(120,22,22,31,15,120)

unique_vals <- unique(data)
print(unique_vals)
solution <- sqrt(49)
print(solution)
round_down <- floor(3.14)
round_up <- ceiling(3.14)
print(round_up)
print(round_down)



# Summary
# R is a powerful statistical programming language with a large community of data enthusiasts.
# You can calculate arithmetic with R and it will follow the normal order of operations
# Data types allow us to classify different pieces of information.
# You can store that information inside of variables
# You can use conditional statements and operators to introduce logic to your code
# You can call a function in R by using the () and passing in the correct arguments
# R programmers have published lots of useful packages that specialize in different tasks