# Read CSV
---
title: "Introduction to Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r}
# load data frame
artists <- read_csv('artists.csv')
artists
```

# Write from Dataframe to CSV
# write_csv(artists,'new_csv_file.csv')


# Head for row numbers & Summary for row stats
---
title: "Introduction to Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
artists <- read_csv('artists.csv')
```

```{r}
# inspect data frame
head(artists)
summary(artists)
```

# Piping
# Pipe operator, or %>%, helps increase the readability of data frame code by piping the value on its left into the first argument 
# of the function that follows it. For example:

# easier to read than
df %>% 
  head()

# this
head(df)



# Select example (just like SQL select)

---
title: "Introduction to Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
artists <- read_csv('artists.csv')
```

```{r}
# select one column
artist_groups <- artists %>%
  select(group)
artist_groups
```

```{r}
# select multiple columns
group_info <- artists %>%
  select(group,spotify_monthly_listeners,year_founded)
group_info
```


# Select All Columns except some
---
title: "Introduction to Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
artists <- read_csv('artists.csv')
```

```{r}
# select all columns except one
no_albums <- artists %>% 
  select(-albums)
no_albums
```

```{r}
# select all columns except a set
df_cols_removed <- artists %>% 
  select(-genre,-spotify_monthly_listeners,-year_founded)
df_cols_removed
```